---
title: "R Notebook"
output: html_notebook
---
```{r}
library(Seurat)
library(Matrix)

# Step 1: Load count matrix
counts <- load counts 

# Step 2: Load gene names
genes <- load genes 

# Step 3: Load barcodes
barcodes <- load barcodes 
rownames(counts) <- genes
colnames(counts) <- barcodes$TAG



seurat_obj <- CreateSeuratObject(counts = counts)
```


```{r}
# Match metadata to expression matrix
metadata <- barcodes
metadata <- metadata[colnames(counts), ]

# Step 7: Add metadata to Seurat
seurat_obj <- AddMetaData(seurat_obj, metadata)
```




```{r}
seurat_obj <- NormalizeData(seurat_obj, normalization.method = "LogNormalize", scale.factor = 10000)
seurat_obj <- FindVariableFeatures(seurat_obj, selection.method = "vst", nfeatures = 2000)
seurat_obj <- ScaleData(seurat_obj)
seurat_obj <- RunPCA(seurat_obj, features = VariableFeatures(object = seurat_obj))
ElbowPlot(seurat_obj)
```

```{r}
seurat_obj <- FindNeighbors(seurat_obj, dims = 1:20)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.5)
```

```{r}
seurat_obj <- RunUMAP(seurat_obj, dims = 1:20)

# 8. Plot
DimPlot(seurat_obj, reduction = "umap", group.by = "broad.cell.type")  # Or another metadata column
```


```{r}
library(ggplot2)
plot <- DimPlot(seurat_obj, reduction = "umap", label = TRUE, label.size = 4.5) + xlab("UMAP 1") + ylab("UMAP 2") +
    theme(axis.title = element_text(size = 18), legend.text = element_text(size = 18)) + guides(colour = guide_legend(override.aes = list(size = 10)))

ClinicalPhenotypes$sample_number <- gsub("ROS","",ClinicalPhenotypes$Subject)
meta3 <- merge(meta, ClinicalPhenotypes, by.x=c("samplenumber"), by.y=c("sample_number"))
rownames(meta3) <- meta3$barcode2

seurat_obj@meta.data <- meta3
```

```{r}
DimPlot(seurat_obj, reduction = "umap", label = TRUE, pt.size = 0.5, split.by = "pathology.group") + NoLegend()
```

```{r}
seurat_obj@meta.data$Pathology[seurat_obj@meta.data$pathology.group == "early-pathology"] <- paste("Pathology")

seurat_obj@meta.data$Pathology[seurat_obj@meta.data$pathology.group == "late-pathology"] <- paste("Pathology")

seurat_obj@meta.data$Pathology[seurat_obj@meta.data$pathology.group == "no-pathology"] <- paste("No-Pathology")

DimPlot(seurat_obj, reduction = "umap", label = TRUE, pt.size = 0.5, split.by = "Pathology") + NoLegend()
```


```{r}
seurat_obj@meta.data$CellType <- Idents(seurat_obj)

seurat_obj@meta.data$celltype_pathology <- paste(seurat_obj@meta.data$CellType, seurat_obj@meta.data$Pathology, sep = "_")
```

```{r}
DimPlot(seurat_obj, reduction = "umap", label = TRUE, pt.size = 0.5, split.by = "Pathology") + NoLegend()
features_OLG <- admeta_GLIA_OLG$gene
features_OLG <- unique(features_OLG)

features_OPC <- admeta_GLIA_OPC$gene
features_OPC <- unique(features_OPC)

FeaturePlot(seurat_obj, features = c("HHIP"), split.by = "Pathology")

DotPlot(seurat_obj, features = features_OLG, split.by = "Pathology") + RotatedAxis()

oligo_Idents <- c("OLGs_1_Pathology","OLGs_1_No-Pathology","OLGs_2_Pathology","OLGs_2_No-Pathology","OLGs_3_Pathology","OLGs_3_No-Pathology","OLGs_4_Pathology","OLGs_4_No-Pathology","OPC_Pathology","OPC_No-Pathology")
oligo_Idents <- as.factor(oligo_Idents)


```

```{r Differential expression testing}
Idents(seurat_obj) <- paste("x")
seurat_obj$celltype.stim <- paste(seurat_obj$celltype_pathology)
Idents(seurat_obj) <- "celltype.stim"



OLG1.de.markers <- FindMarkers(seurat_obj, ident.1 = "OLGs_1_Pathology", ident.2 = "OLGs_1_No-Pathology")


OLG2.de.markers <- FindMarkers(seurat_obj, ident.1 = "OLGs_2_Pathology", ident.2 = "OLGs_2_No-Pathology")


OLG3.de.markers <- FindMarkers(seurat_obj, ident.1 = "OLGs_3_Pathology", ident.2 = "OLGs_3_No-Pathology")


OLG4.de.markers <- FindMarkers(seurat_obj, ident.1 = "OLGs_4_Pathology", ident.2 = "OLGs_4_No-Pathology")


OPC.de.markers <- FindMarkers(seurat_obj, ident.1 = "OPC_Pathology", ident.2 = "OPC_No-Pathology")


