---
title: "WGCNA Analysis Pipeline"
author: "Your Name"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction
This R Markdown document provides a workflow for performing Weighted Gene Co-expression Network Analysis (WGCNA), module–trait correlations, EWCE cell-type enrichment and hub gene identification. 

# Load Required Packages
```{r}
library(WGCNA)
library(GO.db)
library(Biobase)
library(impute)
library(preprocessCore)
library(EWCE)
library(ewceData)
library(SingleCellExperiment)
library(ggplot2)
options(stringsAsFactors = FALSE)
```

# 1. Data Input and Preprocessing
## Load expression matrix and trait data
```{r}
datExpr <- read.csv("expression_matrix.csv", row.names = 1)
traitData <- read.csv("trait_data.csv", row.names = 1)
```

## Sample clustering to identify outliers
```{r}
sampleTree <- hclust(dist(datExpr), method = "average")
plot(sampleTree, main = "Sample Clustering", xlab = "", sub = "")
abline(h = 37000, col = "red")
```

## Remove outlier samples
```{r}
clust <- cutreeStatic(sampleTree, cutHeight = 37000, minSize = 10)
keepSamples <- (clust == 1)
datExpr <- datExpr[keepSamples, ]
traitData <- traitData[rownames(datExpr), ]
```

# 2. Data Transformation & Filtering
## Convert to numeric and log-transform
```{r}
datExpr <- as.data.frame(sapply(datExpr, as.numeric))
datExpr <- log2(datExpr + 1)
```

## Filter lowly expressed genes
```{r}
mean_expression <- colMeans(datExpr)
filtered_data <- datExpr[, mean_expression >= 1.5]
```

# 3. Soft-threshold Selection
```{r}
powers <- c(1:10, seq(12, 20, 2))
sft <- pickSoftThreshold(filtered_data, powerVector = powers, verbose = 5)
```

# 4. Network Construction and Module Detection
```{r}
bwnet <- blockwiseModules(
  filtered_data,
  power = 12,
  networkType = "signed",
  minModuleSize = 200,
  deepSplit = 2,
  mergeCutHeight = 0.3,
  numericLabels = TRUE,
  saveTOMs = TRUE,
  saveTOMFileBase = "TOM-blockwise",
  verbose = 5
)
```

# 5. Module–Trait Relationships
```{r}
moduleColors <- labels2colors(bwnet$colors)
MEs <- orderMEs(bwnet$MEs)
moduleTraitCor <- cor(MEs, traitData, use = "pairwise.complete.obs")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nrow(filtered_data))
```

## Plot heatmap
```{r}
labeledHeatmap(
  Matrix = moduleTraitCor,
  xLabels = names(traitData),
  yLabels = names(MEs),
  colors = blueWhiteRed(50),
  main = "Module–Trait Relationships"
)
```

# 6. EWCE: Cell-type Enrichment
```{r}
ctd <- ewceData::ctd()
moduleGenesList <- probes
results <- bootstrap_enrichment_test(
  sct_data = ctd,
  hits = moduleGenesList,
  genelistSpecies = "mouse",
  sctSpecies = "mouse",
  reps = 1000,
  no_cores = 2,
  annotLevel = 2
)
```

# 7. EWCE Results Plot
```{r}
EWCE::ewce_plot(
  total_res = results$results,
  mtc_method = "BH"
)
```

# 8. Hub Gene Identification for 'pink' module
```{r}
geneModuleMembership <- cor(filtered_data, MEs, use = "p")
module <- "pink"
moduleGenes <- moduleColors == module
kME <- geneModuleMembership[moduleGenes, paste0("ME", module)]
hubGenes <- rownames(kME)[order(kME, decreasing = TRUE)]
topHubGenes <- hubGenes[1:10]
```

# 9. Export to Cytoscape
```{r}
TOM <- TOMsimilarityFromExpr(filtered_data[, moduleGenes], power = 12)
probes <- colnames(filtered_data)[moduleGenes]
dimnames(TOM) <- list(probes, probes)
exportNetworkToCytoscape(
  TOM,
  edgeFile = paste0("Cytoscape_edges_", module, ".txt"),
  nodeFile = paste0("Cytoscape_nodes_", module, ".txt"),
  weighted = TRUE,
  threshold = 0.02
)
```
